---
title: "Exploring R & RStudio"
date: "`r Sys.Date()`"
categories: R
tags:
- R
- RStudio
output:
  blogdown::html_page:
    toc: true
    number_sections: true
type: book
weight: 30
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# Base packages
One of R’s attractive features is its rich collection of packages designed for specific applications and techniques. 

Packages allow users to share R functions and data with other users. 

Some packages come already installed with R, others must be downloaded separately from a CRAN repository or other locations such as GitHub or personal websites.



# Installing Packages from CRAN
There are thousands of additional R packages to choose from. Most can be accessed from the CRAN repository. 

In order to use a package, it needs to be installed on your computer by running `install.packages("name_of_package")` (do not forget `""` around the name of the package).


In the following example, the library ggplot2 is installed from CRAN.

```{r}
install.packages("ggplot2")
```

The installation is usually straightforward and if any other packages need to be installed, RStudio will install those as well as long as the Install dependencies option is checked.

# Loading Packages 
Once the package is installed, or is a base package, you must load the package and only after it has been loaded you can use all the functions and datasets it contains. To load a package, run `library(name_of_package)` (this time `""` around the name of the package are optional).

In the following example, the library ggplot2 is loaded.

```{r}
library(ggplot2)
```

