---
title: "Faculty Data Set"
date: "`r Sys.Date()`"
categories: R
tags:
- Visualización
- ggplot
output:
  blogdown::html_page:
    toc: true
type: book
weight: 20
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

## Context

Evaluating university faculty teaching and research performance is necessary for multiple reasons, including assurance of student learning and informing administrative decision-making.

In this workshop, we will use a simulated dataset consisting of faculty employees information. The data set will be use for learning data visualization in R. The dataset is available in ´xlsx´ and you can download it [here](https://www.dropbox.com/s/rsee0jk8v205jba/FacultyAffairs.xlsx?dl=0).

## Content

The dataset contains 20 variables and 311 rows of data (number of faculty employees). It reports data metrics related to faculty´s demographics, salary, teaching and research performance, among others.

```{r, echo=FALSE}

library(kableExtra)

Dimension <- c("Human Resources","Human Resources", "Human Resources", "Teaching","Teaching","Teaching","Teaching","Research","Research","Service","Service")

Element <- c("Demographic information","Personal information","Employmnet information", "Effectiveness of course delivery",
"Quality of course content","
Effectiveness in Managing Multiple Courses",
"Effectiveness in Mentoring Students",
"Quality of Research",
"Productivity in Research",
"Service to the Institution",
"Service to Society")

Description <- c("Demographic characteristics of the faculty members","Personal information of the faculty members","Information related to appointments of faculty members", "The extent to which a faculty member can lecture, lead discussion, respond to student questions, evaluate assignments and use instructional technology",
"The extent to which knowledge presented is valid. Course content includes: course readings, instructor-developed resources, class exercises, clinical practice, and experiments",
"Ability to manage multuple courses",
"Ability to encourage students and to help them understand the factors that contribute to academic success",
"Evidence of making a substantive contribution to one or more fields of study",
"Capacity to conduct research that is of a high enough quality to regularly appear in refereed publications and be regularly included in conference presentations",
"Involvement in university committees",
"Engaging in community outreach or community engagement effort")

```

```{r, echo=FALSE}

datadic<-data.frame(Element,Description)
kable(datadic,booktabs = T) %>% 
  row_spec(0,font_size = 22) %>% 
  pack_rows(index=c("Human Resources " = 3, "Teaching" = 4, "Research" = 2, "Service" = 2))
```

## Data Dictionary

```{r, echo=FALSE}

library(kableExtra)

Variable <- c("EmployeeName","Salary","Sex","MaritalDesc","DaysLateLast30", "Absences","Age","YearBirth","Nationality","School","Course","Appointment","Satisfaction","ResearchBudget","FirstYear","Articles","Students","Evaluation","Community", "ResearchArea")


Description <- c("Faculty’s full name", "The faculty´s yearly salary. $ U.S. Dollars", " 	Sex - M or F", " The marital status of the faculty (divorced, single, widowed, separated, etc)", " The number of times that the faculty was late to work during the last 30 days" ," No. of times the faculty was absent from work during the year", "Age", "Faculty's year of birth", "Faculty's nationality","School of faculty´s appointment", "No. of courses instructed during the year", "Appointment ranks - 1 = Assistant Professor, 2 = Associate Professor, 3 = Professor", "Results from the last labour satisfaction survey - Value between 0 to 5", "Annual Research Budget", "Year of firts appointment", "No.of articles per year published in peer-reviewed journals during the year", "No.of supervised theses during the year", "Results from the last Teacher evaluation survey - Value between 0 to 5","No.of outreach activities during the year", "Faculty’s research fueld")
                 
Datatype <- c("Character","Numeric","Character","Character","Numeric","Numeric","Numeric","Numeric","Character","Character","Numeric","Numeric","Numeric","Character","Numeric","Numeric","Numeric","Numeric","Numeric","Character")


datadic<-data.frame(Variable,Datatype,Description)
kable(datadic)
```
