---
title: "Day 1: Our Dataset"
date: "`r Sys.Date()`"
categories: R
tags:
- Visualización
- ggplot
output:
  blogdown::html_page:
    toc: true
    number_sections: true

type: book
weight: 40
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE)
```

# Context

Evaluating university faculty teaching and research performance is necessary for multiple reasons, including assurance of student learning and informing administrative decision-making.

In this workshop, we will use a simulated dataset consisting of faculty employees information. The data set will be use for learning data visualization in R. The dataset is available in ´xlsx´ and you can download it [here](https://www.dropbox.com/s/rsee0jk8v205jba/FacultyAffairs.xlsx?dl=0).

# Content

The dataset contains 32 variables and 850 rows of data. It reports data metrics related to four different dimensions: human resources, teaching performance, research performance and services.

```{r, echo=FALSE}
Dimension <- c("Human Resources", "Human Resources", "Teaching","Teaching","Teaching","Research","Service","Service")

Element <- c("Personal information","Employmnet information", "Effectiveness of course delivery","
Effectiveness in Managing Multiple Courses",
"Effectiveness in Mentoring Students",
"Productivity in Research",
"Service to the Institution",
"Service to Society")

Description <- c("Personal information of the faculty members","Information related to appointments of faculty members",
"The extent to which knowledge presented is valid. Course content includes: course readings, instructor-developed resources, class exercises, clinical practice, and experiments",
"Ability to manage multuple courses",
"Ability to encourage students and to help them understand the factors that contribute to academic success",
"Capacity to conduct research that is of a high enough quality to regularly appear in refereed publications and be regularly included in conference presentations",
"Involvement in university committees",
"Engaging in community outreach or community engagement effort")

```

```{r, echo=FALSE}
library(kableExtra)

datadesc<-data.frame(Element,Description)

kable(datadesc,booktabs = T) %>% 
  kable_styling(font_size = 16) %>% 
  row_spec(0,font_size = 22) %>% 
  pack_rows(index=c("Human Resources " = 2, "Teaching" = 3, "Research" = 1, "Service" = 2))
```

# Data Dictionary
Variables included in th dataset are defined in the data dictionary. 

```{r, echo=FALSE}
Variable <- c("AcadYear",
              "EmployeeName",
              "Sex",
              "YearBirth",
              "Nationality",
              "Appointment",
              "SchoolApp",
              "FirstYearApp",
              "Salary",
              "DaysLateLast30",
              "Absences",
              "JobSatisfaction",
              "SelfEvalResp",
              "SelfEvalTech",
              "SelfEvalInnov",
              "StudentEvalResp",
              "StudentEvalTech",
              "StudentEvalInnov",
              "CoursesUnd",
              "CoursesGrad",
              "MentoringResProj",
              "MentoringTheses",
              "ResearchArea",
              "NumPublication",
              "SubManuscripts",
              "NumCitations",
              "NumBooks",
              "NumConferences",
              "ResGrantsAwa",
              "ResGrantsSub",
              "NumComUniv",
              "NumOut")

              
Description <- c("Academic year",
                 "Faculty’s full name",
                 "Sex (M or F)",
                 "Faculty's year of birth",
                 "Faculty's nationality",
                 "Appointment ranks (1 = Assistant Professor, 2 = Associate Professor, 3 = Professor)",
                 "School of  appointment",
                 "Year of first appointment",
                 "The faculty´s yearly salary ($U.S. Dollars)", 
                 "No. of times the faculty was late to work during the year",
                 "No. of times the faculty was absent from work during the year",            
                 "Results from the last labour satisfaction survey using a 5-point Likert scale (from 1 extremely dissatisfied to 5 extremely satisfied)",
                  "Results from the last self-assestment survey about spped of response to students using a 5-point Likert scale (from 1 extremely slow to 5 extremely fast)",
                  "Results from the last self-assestment survey about the use of instructional technology using a 5-point Likert scale (from 1 never to 5 always)",
                  "Results from the last self-assestment survey about the use of educational innovation approaches using a 5-point Likert scale (from 1 never to 5 always)",
                "Results from the last evaluation by students about spped of response to students using a 5-point Likert scale (from 1 extremely slow to 5 extremely fast)",
                "Results from the last evaluation by students about the use of instructional technology using a 5-point Likert scale (from 1 never to 5 always)",
                "Results from the last evaluation by students about the use of educational innovation approaches using a 5-point Likert scale (from 1 never to 5 always)",
                "No. of undergraduate courses instructed during the year",
                "No. of graduate courses instructed during the year",
                "No. of students mentored in research during the year",
                "No. of theses supervised during the year", 
                "Faculty’s research area",
                "No. of articles published in peer-reviewed journals during the year",
                "No. of articles submitted to peer-reviewed journals during the year",
                "Faculty´s cummulative number of citations",
                "No. of books or book chapters published during the year",
                "No.of international conferences presentations during the year",
                "No. of research grants submitted",
                "No. of research grants awarded",
                "No. of  of participations in institutional committee",
                "No. of participations in outreach activities")
                
            
Element <- c( " ",
              "Personal information",
              "Personal information",
              "Personal information",
              "Personal information",
              "Employment information",
              "Employment information",
              "Employment information",
              "Employment information",
              "Employment information",
              "Employment information",
              "Employment information",
              "Effectiveness of course delivery",
              "Effectiveness of course delivery",
              "Effectiveness of course delivery",
              "Effectiveness of course delivery",
              "Effectiveness of course delivery",
              "Effectiveness of course delivery",
              "Effectiveness in managing multiple courses",
              "Effectiveness in managing multiple courses",
              "Effectiveness in mentoring students",
              "Effectiveness in mentoring students",
              "Productivity in research",
              "Productivity in research",
              "Productivity in research",
              "Productivity in research",
              "Productivity in research",
              "Productivity in research",
              "Productivity in research",
              "Productivity in research",
              "Service to the institution",
              "Service to the society")

```

```{r , echo=FALSE}
library(kableExtra)
datadic<-data.frame(Variable,Description,Element)
kable(datadic) %>%
  kable_styling(font_size = 16) %>% 
  row_spec(0,font_size = 22)

```

# Visualizations in a glace

I have developed An application programming interface (API) to explore the visualizations that we will explore in this workshop. To do that, follow these steps:

1.   Download the data set [here](https://www.dropbox.com/s/rsee0jk8v205jba/FacultyAffairs.xlsx?dl=0).

2. Upload the file.

 <iframe src="https://pacortes.shinyapps.io/FacultyReport/" width=1000 height=800"></iframe>



